### 2021-06-11 paper/paper.tex
Transfered over files from original directories. Will now track everything to do with this project through git.

\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}

\title{D-optimal Designs for Logistic Regression using a Genetic Algorithm}
\author{Will Gertsch}

\usepackage{setspace}
\doublespacing

\begin{document}
\maketitle

\begin{abstract}
Logistic regression is a popular model for a binary outcome. While locally D-optimal designs can be found analytically for the case of a linear predictor, adding a quadratic term or constricting the design interval makes finding the optimal design more challenging. In most of these cases, there is no known analytical result. We use  a genetic algorithm to find locally D-optimal designs for linear and quadratic models on any design interval. We are able to confirm previous results and find new designs in several cases. 
\end{abstract}

\section{Introduction}
\section{Preliminaries}
\subsection{Logistic Regression}
Suppose we have a binary outcome modeled as $y_i \sim \text{Bernoulli}(p_i)$ for observations $i = 1, \dots, n$ and $0 \leq p_i < 1$. We can add a predictor $x_i$ into the model using the logit link function
$$
\log\left( \frac{p_i}{1-p_i}\right) = \eta_i
$$
where $\eta_i = \beta_0  + \beta_1 x_{i}$ or $\eta_i = \beta_0  + \beta_1 x_{i} + \beta_2 x_i^2$. Rearranging, we obtain an expression for the mean of $y_i$ given $\eta_i$
$$
p_i = \frac{e^{\eta_i}}{1+e^{\eta_i}}= \frac{1}{1+e^{-\eta_i}} 
$$

The information matrix for this model may be expressed as
$$
M(\beta) = \sum_{i=1}^k p_i (1-p_i) f(x_i) f(x_i)' = X'WX
$$
where $W$ is a diagonal weight matrix with entries $p_i (1-p_i)$. The design matrix $X$ has rows $f(x_i)' = (1, x_i)$ if $\eta_i$ is linear. If $\eta_i$ is quadratic, then $f(x_i)' = (1,x_i,x_i^2)$.
\subsection{Optimal Design}
The goal of optimal design is to find the optimal predictor values $x_1,\dots, x_k $ given a fixed number of observations $N$. Each design point $x_i$ is assigned $n_i$ observations such that $\sum_{i=1}^k n_i = N$. Alternatively, we may assign a weight $w_i$ to each $x_i$ to convey the same information. In this case, $\sum_{i=1}^k w_i = 1$ where $w_i = n_i/N$.

\par A design using weights be written as
$$
\xi =
\begin{pmatrix}
x_1 & \dots & x_k\\
w_1 & \dots & w_k
\end{pmatrix}
$$

The design $x_i$ implies a corresponding information matrix for the logistic regression model. Let
$$
M = M(\xi, \beta) = \sum_{i=1}^k w_ip_i (1-p_i) f(x_i) f(x_i)' = X'WX
$$
be the design and model implied information matrix where $W$ now has diagonal entries $w_i p_i (1-p_i)$.

The optimality of a particular design is determined by some function of the information matrix. It can be shown that the determinant of the information matrix is inversely proportional to the size of the confidence region for the parameters $\beta_0$, $\beta_1$, and $\beta_2$. Therefore, we may define a design as being optimal if it maximizes the determinant of the information matrix. This criterion is called D-optimality. Equivalently, we may minimize
$$
\Psi(M) = -\log(|M|)
$$

It can be shown that $\Psi(M)$ is a convex function of the design points $x_1, \dots, x_k$ and  the weights $w_1, \dots, w_k$. However, $\Psi(M)$ also depends on the parameters $\beta_0$, $\beta_1$, and $\beta_2$. This means that nominal values will need to be supplied for these parameters. Therefore, we may obtain a locally D-optimal design by first providing values for the parameters and then minimizing $\Psi(M)$.

To check if a design is locally D-optimal, we calculate the sensitivity function
$$
ch(x) = g(x) f(x)'M(\xi, \beta)^{-1} f(x) - p
$$
where $p$ is the number of parameters in the model and
$$
g(x) = \frac{\exp(\eta)}{(1+\exp(\eta))^2}
$$
The equivalence theorem says a design is locally D-optimal if and only if $ch(x) \leq 0 $ for all $x$ in the design interval with equality at the design points.

\subsection{Particle Swarm}
\subsection{Genetic Algorithm}
Finding a locally D-optimal design for the logistic model requires we optimize a function of several variables. This can be accomplished analytically for a linear $\eta$, but the complexity dramatically increases for a quadratic $\eta$ and requires a numerical solution. The objective function is complicated and changes between linear and quadratic models. Therefore, we should use an optimization algorithm that is gradient free. Additionally, we should use an algorithm that can incorporate constraints on the design space and require that the weight variables sum to 1.

The genetic algorithm (Holland, 1975) is a nature-inspired metaheuristic optimization algorithm. The algorithm mimics natural selection with the intended goal of ``evolving" a population of solution vectors towards an optimal value. Like other nature-inspired optimization algorithms, the genetic algorithm has the advantages of being gradient free, able to accommodate constraints, and able to escape local optimums.

There are many variations on the genetic algorithm, but these algorithms usually contain the following three steps that are run every iteration of the algorithm. First, the \textit{crossover} operation combines two parent solutions to make two child solutions. Second, the \textit{mutation} operation randomly perturbs the child solutions. Finally, \textit{fitness} is assessed and the solution vectors that have the best objective function values move on to the next generation.

The following paragraphs detail the implementation of the genetic algorithm in PlatEMO. Let $N$ be the population size and let $D$ be the number of variables in the optimization problem.

\begin{flushleft}
\textbf{Selection}
\end{flushleft}
The first step in the genetiec algorithm is to select the mating pool via tournament selection. $N$ paris are sampled with replacement from the population. The objective function values for eahc of these pairs is combared and the most opitmal of the pair is selected.

\begin{flushleft}
\textbf{Crossover}
\end{flushleft}
Once the mating pool has been selected, the two parents enter the crossover stage where genetic information is exchanged. This is done using simulated binary crossover (Deb et. al. 2007).

\begin{flushleft}
\textbf{Mutation}
\end{flushleft}

\begin{flushleft}
\textbf{Fitness}
\end{flushleft}

\section{Results}
The genetic algorithm was implemented using the PlatEMO framework in MATLAB. We used the built-in algorithm ``GA'' with default parameters ($p_C = 1, d_C = 20, p_M = 1, d_M = 20$). We found that running the algorithm for 100000 function evaluations with a swarm size of 1000 was able produce close to optimal designs for every design we tried.

When the optimal number of design points was unknown, we ran the algorithm once with 10 design points and then made a guess at the correct number of design points. The optimal number of design points was easy to find after seeing the sensitivity plot.

\subsection{Linear $\eta$ on $(-\infty, \infty)$}
If $\eta_i = \beta_0 + \beta_1 x_i$ and the design interval is $(-\infty, \infty)$, it can be shown that the locally D-optimal design is equally weighted at $x_1$ and $x_2$ where
$$
x_1, x_2 = \frac{\pm 1.5434 - \beta_0}{\beta_1}
$$

For example, if $\beta_0 = 1$ and $\beta_1 = 1$ then the optimal design is
$$
\xi = \begin{pmatrix}
-1.5434 & 1.5434\\ 0.5 & 0.5
\end{pmatrix}
$$

The genetic algorithm was able to replicate this theoretical result. Table 1 shows designs for several nominal values of $\beta$ with the theoretical designs and the designs generated from the genetic algorithm. The algorithm is able to come very close to the theoretical result. Sensitivity plots confirm that the designs generated by the genetic algorithm are almost indistinguishable from the true optimum. Figure 1 shows one such sensitivity plot.
\begin{table}[ht]
\centering
\begin{tabular}{|c|c|c|}
\hline 
$\mathbf{\beta}$ & \textbf{Theoretical} & \textbf{GA} \\ 
\hline 
0,1 & $\begin{bmatrix}
-1.5434 & 1.5434\\ 0.5 & 0.5
\end{bmatrix}$  & $\begin{bmatrix}
-1.5452 & 1.5435\\ 0.5000 & 0.5000
\end{bmatrix}$ \\ 
\hline 
0.3,0.4 & $\begin{bmatrix}
-4.6085 & 3.1085\\ 0.5 & 0.5
\end{bmatrix}$  & $\begin{bmatrix}
-4.6091 & 3.1083\\ 0.5000 & 0.5000
\end{bmatrix}$ \\ 
\hline 
2,-5 & $\begin{bmatrix}
0.0913 & 0.7087\\ 0.5 & 0.5
\end{bmatrix}$  & $\begin{bmatrix}
0.0923 & 0.7109\\ 0.5000 & 0.5000
\end{bmatrix}$  \\ 
\hline
-4, 10 & $\begin{bmatrix}
0.2457 & 0.5543\\ 0.5 & 0.5
\end{bmatrix}$  & $\begin{bmatrix}
0.2475 & 0.5587\\ 0.5000 & 0.5000
\end{bmatrix}$\\
\hline

\end{tabular} 
\caption{Locally D-optimal designs generated by the genetic algorithm compared to analytic solution.}
\end{table}

\begin{figure}
\centering
\includegraphics[scale=0.5]{figures/fig1.jpg}
\caption{Sensitivity function for design produced by genetic algorithm for a nominal $\beta$.}
\end{figure}

\subsection{Linear $\eta$ on other intervals}
The genetic algorithm can find designs on an arbitrary design interval $[a,b]$. In general, the optimal design will change from the design on $(-\infty, \infty)$ if any of the designs points is no longer included in the interval. Therefore, at least one of the two design points will likely be one of the bounds of the interval. Theoretical justifications for this phenomenon can be found in Sebastiani.

Table 2 shows how the design interval affects the optimal design points. For example, if the interval excludes the global optimal point -1.543, then the optimal design will have one point at the bound of the design interval closest to that point. The other design point is also changed from $1.543$ to 1.796 to adjust for the change in the first point. All the designs remain equally weighted.

\begin{table}
\centering
\begin{tabular}{|c|c|c|c|c|c|}
\hline 
  & $[-\infty, \infty]$  & $[-1, \infty]$ & $[-\infty,1]$ & $[-1,1]$ & $[10, 20]$\\ 
\hline 
$x_1 (T)$ & -1.543 & -1 & -1.796 & -1  & ?\\ 
\hline 
$x_2 (T)$ & 1.543 & 1.796 & 1 & 1 & ?\\

\hline
$x_1 (GA)$ & -1.545 & -1 & -1.796 & -1 & $10$\\
\hline
$x_2 (GA)$ & 1.544 & 1.796 & 1 & 1 & $12$\\
\hline
\end{tabular}
\caption{Design points for locally D-optimal designs on several intervals for $\beta = (0,1)$ from theoretical and genetic algorithm approaches. ?: theoretical result not known.}
\end{table}

For the case when $a,b$ were both greater than the optimal design points on $(-\infty, \infty)$, there was no analytical result. However, the genetic algorithm was able to find an optimal design in this case. The last column of Table 2 shows optimal design on the interval $[10,20]$ and the sensitivity plot in Figure 2 confirms optimality.

\begin{figure}
\centering
\includegraphics[scale=0.5]{figures/fig2.jpg}
\caption{Sensitivity plot for a design generated by the genetic algorithm on the design interval [10, 20].}
\end{figure}

\subsection{Quadratic $\eta$ on $(-\infty, \infty)$}
There are no analytic results for locally D-optimal designs when $\eta_i = \beta_0 + \beta_1 x_i + \beta_2 x_i^2$. However, Fornious was able to find locally D-optimal designs numerically for several nominal $\beta$. If $\beta_1 \neq 0$, then the design shifts along the design space. Therefore, the example nominal values have $\beta_1$ set at 0.

Table 3 compares the results from Fornious with the results from the genetic algorithm. The designs produced by the genetic algorithm are very close to the designs from Fornious and have sensitivity plots that indicate optimality. Some sets of nominal parameter values result in an optimal 3 point design while others have an optimal design with 4 points. Figure 3 shows the sensitivity plot for a 4 point design with $\beta = (2, 0, -0.1)$. Figure 4 shows a 3 point design for $\beta = (-2, 0, -4)$.
\begin{table}
\begin{flushleft}
\begin{tabular}{|c|c|c|}
\hline 
 Nominal values & Fornious & GA \\ 
\hline 
 $2,0,-0.1$ & 
$\begin{bmatrix}
-5.7185 & -2.7017  & 2.7017 & 5.7185\\
0.3138 & 0.1862 & 0.1862 & 0.3138
\end{bmatrix}$  & $\begin{bmatrix}
-5.7229 & -2.6945 & 2.7579 & 5.7229\\
0.3141 & 0.1894 & 0.1840 & 0.3126
\end{bmatrix}$\\ 
\hline 
 $2,0,-4$ & $\begin{bmatrix}
-0.9042 & -0.4272  & 0.4272 & 0.9042\\
0.3138 & 0.1862 & 0.1862 & 0.3138
\end{bmatrix}$  & $\begin{bmatrix}
-0.9051 & -0.4275  & 0.4272 & 0.9038\\
0.3139 & 0.1873 & 0.1850 & 0.3139
\end{bmatrix}$\\ 
\hline 
 $-2,0,-0.1$ & 
$\begin{bmatrix}
-3.9819 & 0.0000  & 3.9819\\
0.3333 & 0.3333 & 0.3333
\end{bmatrix}$   & $\begin{bmatrix}
-3.9818 & -0.0001  & 3.9813\\
0.3333 & 0.3333 & 0.3333
\end{bmatrix}$ \\ 
\hline 
 $-2,0,-4$ & 
$\begin{bmatrix}
-0.6296 & 0.0000  & 0.6296\\
0.3333 & 0.3333 & 0.3333
\end{bmatrix}$ & $\begin{bmatrix}[
-0.6295 & 0.0000  & 0.6296\\
0.3333 & 0.3333 & 0.3333
\end{bmatrix}$\\ 
\hline 
\end{tabular} 
\end{flushleft}

\caption{Designs for a quadratic $\eta$ from Fornious and the genetic algorithm.}
\end{table}

\begin{figure}
\centering
\includegraphics[scale=0.5]{figures/fig3.jpg}
\caption{Sensitivity plot for a 4 point design found for quadratic $\eta$ using the genetic algorithm.}
\end{figure}

\begin{figure}
\centering
\includegraphics[scale=0.5]{figures/fig4.jpg}
\caption{Sensitivity plot for a 3 point design found for quadratic $\eta$ using the genetic algorithm.}
\end{figure}

The locally D-optimal designs for a quadratic $\eta$ have a few interesting features. First all designs are symmetric on the design space. This is easy to see when $\beta_1 = 0$, but also holds true when $\beta_1 \neq 0$. Secondly, the 3 point designs have equal weights at each design point. Finally,  the weights on each side of the symmetric 4 point designs sum to 1/2.

\section{Quadratic $\eta$ on other intervals}
There are many possible ways to restrict the design interval for the quadratic model. This is especially true when the locally D-optimal design on the unbounded interval has 4 points. To simplify the infinite number of nominal value and design interval combinations, we will restrict our focus on the nominal $\beta$ values of $(2, 0, -4)$ and $(-2, 0, -4)$ which give 4 and 3 point designs respectively. By modifying the design interval to include or not include the original optimal design points, we obtain insights that should generalize to other nominal values.

Recall that the locally D-optimal design for $\beta = (2,0,-4)$ on $(-\infty, \infty)$ has design points at $x_1=-0.9051, x_2=-0.4275, x_3=0.4272,$ and  $x_4=0.9038$. If the design interval is constructed such that $x_1$ and $x_4$ are not included, we get a 3 point equally weighted design with points at the bounds of the interval and at the axis of symmetry. Figure 5 shows the design when the interval is $[-0.5, 0.5]$. Restricting the interval to not include any of the original design points and placing the interval to be between $x_2$ and $x_3$ produces a similar design in Figure 6. We may also construct designs where the interval includes $x_3$ and $x_4$.  For the interval $[0,1]$, this produced an equally weighted 3 point design as can bee seen in figure 7.

\begin{figure}
\centering
\includegraphics[scale=0.5]{figures/4pt_include2mid.jpg}
\caption{Sensitivity plot for a design on $[-0.5, 0.5]$ for a quadratic $\eta$.}
\end{figure}

\begin{figure}
\centering
\includegraphics[scale=0.5]{figures/4pt_nonemid.jpg}
\caption{Sensitivity plot for a design on $[-0.3, 0.3]$ for a quadratic $\eta$.}
\end{figure}

\begin{figure}
\centering
\includegraphics[scale=0.5]{figures/4pt_includeend2.jpg}
\caption{Sensitivity plot for a design on $[0,1]$ for a quadratic $\eta$.}
\end{figure}

For the originally 3 point design with $\beta = (-2, 0, -4)$, the results are very similar. Recall the optimal design for these nominal values was equally weighted at $x_1=-0.6295$, $x_2 = 0$,  and $x_3 = 0.6296$. If $x_2, x_3$ are included in the design interval, the result is an equally weighted 3 point design. Figure 8 shows the optimal design on the interval $[0, 1]$.

\begin{figure}
\centering
\includegraphics[scale=0.5]{figures/3pt_2include.jpg}
\caption{Sensitivity plot for a design on $[0,1]$ for a quadratic $\eta$.}
\end{figure}

Unfortunately, not all designs have a sensitivity plot that can easily plotted. The genetic algorithm still finds the design but it is difficult to judge from the plot whether the design is truly optimal. For example, Figure 9 shows the plot for $\beta = (-2, 0, -4)$ when the design interval is $[1,2]$. The sensitivity function at the design points is close to zero, but most of the curve is also very close to zero.

\begin{figure}
\centering
\includegraphics[scale=0.5]{figures/3pt_noneend.jpg}
\caption{Sensitivity plot for a design on $[1,2]$ for a quadratic $\eta$.}
\end{figure}

\section{Applications}
\section{Conclusions}
\section{References}
\section{Appendix}
\end{document}

### 2021-06-11 presentation/presentation.tex
Transfered over files from original directories. Will now track everything to do with this project through git.

\documentclass[11pt]{beamer}
\usetheme{CambridgeUS}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\author{Will Gertsch}
\title{D-optimal Designs for Logistic Models using Metaheuristics}
%\setbeamercovered{transparent} 
%\setbeamertemplate{navigation symbols}{} 
%\logo{} 
%\institute{} 
%\date{} 
%\subject{} 
\begin{document}

\begin{frame}
\titlepage
\end{frame}

%\begin{frame}
%\tableofcontents
%\end{frame}

\begin{frame}{Outline}
\begin{enumerate}
\item Logistic regression model and information matrix
\item Optimal design background
\item Particle swarm and genetic algorithm
\item D-optimal designs for a linear predictor
\item D-optimal designs for a quadratic predictor
\end{enumerate}
\end{frame}

\begin{frame}{Logistic Model}
Suppose $y_i \sim \text{Bernoulli}(p_i)$ for $i = 1, \dots, n$. The logit link function is defined as
$$
p_i = \frac{1}{1+e^{-\eta_i}} = \frac{e^{\eta_i}}{1+e^{\eta_i}}
$$
where $\eta_i = \beta_0  + \beta_1 x_{i}$
and $0 \leq p_i < 1$

We may also use a quadratic model $\eta_i = \beta_0  + \beta_1 x_{i} + \beta_2 x_i^2$.
\end{frame}

\begin{frame}{Information Matrix}
The information matrix for the logistic model can be derived as 
$$
M(\beta) = \sum_{i=1}^k p_i (1-p_i) f(x_i) f(x_i)' = X'WX
$$
where $W$ is a diagonal weight matrix with entries $p_i (1-p_i)$. The design matrix $X$ has rows $f(x_i)' = (1, x_i)$ if $\eta_i$ is linear. If $\eta_i$ is quadratic, then $f(x_i)' = (1,x_i,x_i^2)$. 
\end{frame}

%\begin{frame}{Information Matrix Example}
%Suppose we have $\eta_i = \beta_0 + \beta_1 x_i$ and $n = 2$. Then the information matrix is
%$$
%I\begin{pmatrix}
%\beta_0\\
%\beta_1
%\end{pmatrix}
%=
%p_1 (1-p_1)
%\begin{pmatrix}
%1 & x_1\\
%x_1 & x_1^2
%\end{pmatrix}
%+
%p_2 (1-p_2)
%\begin{pmatrix}
%1 & x_1\\
%x_1 & x_1^2
%\end{pmatrix}
%$$
%Note that  $I$ depends on $\beta_0$ and $\beta_1$ through $p_1$ and $p_2$.
%
%\end{frame}

\begin{frame}{Design Notation}
We denote a design $\xi$ using weight notation. If we have $N$ samples total and $n_i$ samples for each design point $x_1 \dots, x_k$, let 
$$
\xi =
\begin{pmatrix}
x_1 & \dots & x_k\\
w_1 & \dots & w_k
\end{pmatrix}
$$
where $w_i = n_i/N$ are weights.
\end{frame}

\begin{frame}{Information Matrices}
In optimal design, a design $\xi$  implies a corresponding information matrix
$$
M = M(\xi, \beta) = \sum_{i=1}^k w_ip_i (1-p_i) f(x_i) f(x_i)' = X'WX
$$
where $W$ now has diagonal entries $w_i p_i (1-p_i)$
\end{frame}

\begin{frame}{Locally Optimal Designs}
For a given value of the parameter vector $\beta$, we can find a locally optimal design by minimizing functions of the information matrix. Below are some common examples:
\begin{itemize}
\item \textbf{D-optimality:} $-\log\left(\text{det}(M)\right)$
\item A-optimality: $\text{tr}(M^{-1})$
\item E-optimality: $-\text{max}_{M} \text{min}_{Mv = \lambda v} \lambda$
\end{itemize}

In this presentation, we will focus on D-optimality.
\end{frame}

\begin{frame}{Equivalence Theorem for D-optimality}
We can test if a design is D-optimal by computing the sensitivity function
$$
ch(x) = g(x) f(x)'M(\xi, \beta)^{-1} f(x) - p
$$
where $p$ is the number of parameters in the model and 
$$
g(x) = \frac{\exp(\eta)}{(1+\exp(\eta))^2}
$$
$\eta = \beta_0  + \beta_1 x$ or $\eta = \beta_0 + \beta_1 x + \beta_2 x^2$
\end{frame}

\begin{frame}{Plotting ch(x)}
If $ch(x) = g(x) f(x)'M(\xi, \beta)^{-1} f(x) - p \leq 0$ with equality at the design points, then the design $\xi$ is D-optimal. In other words, we can plot $ch(x)$ to see if a design is optimal.\\

\begin{figure}[h]
\centering
\includegraphics[scale=0.3]{prob3p2.jpg}
\caption{ch(x) for a D-optimal design}

\end{figure}

\end{frame}

\begin{frame}{Review of Previous Results}
Papers on the logistic model D-optimality:
\begin{itemize}
\item Mathew, Sinha (2001) derive D-optimal designs for linear $\eta$.
\item Sebastiani, Settimi (1992) derive designs on different design intervals for linear $\eta$.
\item Lall et al. (2018) use a Fedorov algorithm to find designs on $[-1,1]$.
\item Fornius (2005) found designs numerically for quadratic $\eta$.
\end{itemize}
\end{frame}

\begin{frame}{My Contributions}
I wrote software that uses metaheuristic optimization algorithms to
\begin{itemize}
\item Find D-optimal designs for linear and quadratic logistic models.
\item Find designs on any design interval.
\item Change number of design points.
\end{itemize}
\end{frame}

\begin{frame}{Choice of algorithm}
I used 2 algorithms to find designs:
\begin{itemize}
\item Particle swarm optimization (PSO)
\item Genetic algorithm (GA)
\item Using implementation in PlatEMO with default parameters.
\item PSO may work better than GA or vice-versa
\end{itemize}
These algorithms were chosen because they worked the best out of all the other single-objective algorithms in PlatEMO for this problem.
\end{frame}

\begin{frame}{Particle Swarm Optimization}
Particle swarm optimization (PSO) (Kennedy, Ebernhart 1995) is a nature-inspired algorithm to find the optimal value of an objective function. It is based on fish and bird schooling behavior.\\
\textbf{Algorithm}:
\begin{enumerate}
\item Initialize $N$ particles $x_i$ with velocities $v_i$.
\item Find particle with best objective value.
\item Update velocity $v_i^{t+1}$ for all particles such that they are drawn towards the current best global value and towards the current best for particle $i$.
\item Update location $x_i^{t+1} = x_i^{t} + v_i^{t+1}$ for all particles.
\item Repeat steps 2-4.
\end{enumerate}
%PlatEMO: Velocity calculation includes a constant inertial weight that can be specified by user.
\end{frame}

\begin{frame}{Genetic Algorithm}
The genetic algorithm (GA) (Holland 1975) is based on natural selection.It has 3 main steps: crossover, mutation, and selection.
\begin{itemize}
\item $x_i$'s are encoded as strings called chromosomes.
\item Crossover and mutation steps operate on these chromosomes.
\item Diagrams from \textit{Nature-Inspired Optimization Algorithms} by Xin-She Yang.
\end{itemize}
\end{frame}

\begin{frame}{Crossover}

\centering
\includegraphics[scale=0.5]{crossover.png}

\end{frame}

\begin{frame}{Mutation}
\centering
\includegraphics[scale=0.5]{mutation.png}
\end{frame}

\begin{frame}{Genetic Algorithm}
\textbf{Algorithm}:
\begin{enumerate}
\item Initialize $N$ solutions.
\item Crossover: swap characteristics of 2 parent solutions with prob. $p_c$ to produce 2 child solutions.
\item Mutation: randomly alter a characteristic of a child solutions with prob. $p_m$.
\item Selection: Accept child solutions if fitness increases.
\item Best $N$ solutions go on to next generation.
\item Repeat steps 2-5.
\end{enumerate}
This is the basic format of a genetic algorithm. Details may vary with implementation.
\end{frame}

\begin{frame}{Results from Mathew, Sinha}
Mathew, Sinha show that the locally $D$-optimal design for a logistic model with $\eta = \beta_0 + \beta_1 x$ has two equally weighted design points
$$
x_1, x_2 = \frac{\pm 1.5434 - \beta_0}{\beta_1}
$$
\end{frame}

\begin{frame}{Results from Mathew, Sinha}
Algorithms were run with swarm size of 1000 and for 1000000 objective function evaluations. Both algorithms had the same initial values.
\begin{tabular}{|c|c|c|c|}
\hline 
$\beta$ & Theoretical & PSO & GA \\ 
\hline 
0,1 & $\begin{bmatrix}
-1.5434 & 1.5434\\ 0.5 & 0.5
\end{bmatrix}$ & $\begin{bmatrix}
-1.5434 & 1.5434\\ 0.5000 & 0.5000
\end{bmatrix}$ & $\begin{bmatrix}
-1.5452 & 1.5435\\ 0.5000 & 0.5000
\end{bmatrix}$ \\ 
\hline 
0.3,0.4 & $\begin{bmatrix}
-4.6085 & 3.1085\\ 0.5 & 0.5
\end{bmatrix}$ & $\begin{bmatrix}
-4.6085 & 3.1085\\ 0.5000 & 0.5000
\end{bmatrix}$ & $\begin{bmatrix}
-4.6091 & 3.1083\\ 0.5000 & 0.5000
\end{bmatrix}$ \\ 
\hline 
2,-5 & $\begin{bmatrix}
0.0913 & 0.7087\\ 0.5 & 0.5
\end{bmatrix}$ & $\begin{bmatrix}
0.0913 & 0.7087\\ 0.5000 & 0.5000
\end{bmatrix}$ & $\begin{bmatrix}
0.0923 & 0.7109\\ 0.5000 & 0.5000
\end{bmatrix}$  \\ 
\hline 
\end{tabular}

PSO tends to converge quicker for this problem but GA comes very close to theoretical solution.
\end{frame}

\begin{frame}{Results from Lall et al.}
Lall et. al use a modified Fedorov algorithm to find D-optimal designs on $[-1,1]$. PSO and GA are run with the same parameters as last slide.
\begin{tabular}{|c|c|c|c|c|c|c|c|}
\hline 
 $\beta_0$ & $\beta_1$ & $x_1 (F)$ & $x_2 (F)$ & $x_1 (PSO)$ & $x_2 (PSO)$ & $x_1 (GA)$ & $x_2 (GA)$\\
\hline 
 0.1 & 0.5 & -1 & 1 & -1 & 1 & -1 & 1\\ 
\hline 
 1 & 1 & -1 & 1 & -1 & 1 & -1 & 1\\ 
\hline 
 1 & 4 & -0.636 & 0.136 & -0.636 & 0.136 & -0.636 & 0.136\\ 
\hline 
\end{tabular}
PSO and GA both are able to confirm the solution.
\end{frame}

\begin{frame}{Results Sebastiani}
Sebastiani derives theoretical results for all possible combinations of design intervals. The resulting designs are modifications of the optimal design on $(-\infty, \infty)$. To illustrate, I selected $\beta_0 = 0, \beta_1=1$. All resulting designs were equally weighted.

\begin{tabular}{|c|c|c|c|c|c|}
\hline 
  & $[-\infty, \infty]$  & $[-1, \infty]$ & $[-\infty,1]$ & $[-1,1]$ & $[10, 20]$\\ 
\hline 
$x_1 (T)$ & -1.543 & -1 & -1.796 & -1  & ?\\ 
\hline 
$x_2 (T)$ & 1.543 & 1.796 & 1 & 1 & ?\\ 
\hline 
$x_1 (PSO)$ & -1.543 & -1 & -1.796 & -1 & $10^*$\\
\hline
$x_2 (PSO)$ & 1.543  & 1.796 & 1 & 1 & $10^*$\\
\hline
$x_1 (GA)$ & -1.545 & -1 & -1.796 & -1 & $10$\\
\hline
$x_2 (GA)$ & 1.544 & 1.796 & 1 & 1 & $12$\\
\hline
\end{tabular} \\
$*$ PSO had trouble converging to the correct solution in this case. \\Note also that the last case is not covered in Sebastiani.
\end{frame}

\begin{frame}{Results from Fornius}
Fornious (2005) derives D-optimal designs for the quadratic model with $\eta_i = \beta_0 + \beta_1 x_i + \beta_2 x_i^2$. Designs have either 3 or 4 support points depending on shape of response.

\begin{tabular}{|c|c|}
\hline 
 Nominal values & Design \\ 
\hline 
 $2,0,-0.1$ & 
$\begin{bmatrix}
-5.7185 & -2.7017  & 2.7017 & 5.7185\\
0.3138 & 0.1862 & 0.1862 & 0.3138
\end{bmatrix}$  \\ 
\hline 
 $2,0,-4$ & $\begin{bmatrix}
-0.9042 & -0.4272  & 0.4272 & 0.9042\\
0.3138 & 0.1862 & 0.1862 & 0.3138
\end{bmatrix}$   \\ 
\hline 
 $-2,0,-0.1$ & 
$\begin{bmatrix}
-3.9819 & 0  & 3.9819\\
1/3 & 1/3 & 1/3
\end{bmatrix}$   \\ 
\hline 
 $-2,0,-4$ & 
$\begin{bmatrix}
-0.6296 & 0  & 0.6296\\
1/3 & 1/3 & 1/3
\end{bmatrix}$  \\ 
\hline 
\end{tabular} \\
$\beta_1 \neq 0$ shifts the design.\\ What about on different design intervals?
\end{frame}

% better to just use plots for these designs
%\begin{frame}{Fornius PSO}
%PSO struggles to find the optimal design for the 4 point designs even after many iterations. 3 point designs seem to be easier.
%
%\begin{tabular}{|c|c|}
%\hline 
% Nominal values & Design \\ 
%\hline 
% $2,0,-0.1$ & 
%$\begin{bmatrix}
%-5.8244 & -1.6960  & 3.0160 & 5.5276\\
%0.3629 & 0.3006  & 0.0698 & 0.2668
%\end{bmatrix}$  \\ 
%\hline 
% $2,0,-4$ & $\begin{bmatrix}
%-0.9208 & -0.3372 & 0.7906 & 0.8980\\
%0.3355 & 0.3124 & 0.1645 & 0.1876 
%\end{bmatrix}$   \\ 
%\hline 
% $-2,0,-0.1$ & 
%$\begin{bmatrix}
%-3.9819 & 0  & 3.9819\\
%1/3 & 1/3 & 1/3
%\end{bmatrix}$   \\ 
%\hline 
% $-2,0,-4$ & 
%$\begin{bmatrix}
%-0.6295 & 0.0001  & 0.6297\\
%0.3346 & 0.3331 & 0.3323
%\end{bmatrix}$  \\ 
%\hline 
%\end{tabular} 
%\end{frame}
\begin{frame}{PSO Results 1}
PSO can find optimal designs for the 3 point cases.
\includegraphics[scale=0.28]{quadplots/PSO_3.jpg}
\includegraphics[scale=0.28]{quadplots/PSO_4.jpg}
\end{frame}

\begin{frame}{PSO Results 2}
PSO has trouble finding the optimal design when the design has 4 points.
\includegraphics[scale=0.28]{quadplots/PSO_1.jpg}
\includegraphics[scale=0.28]{quadplots/PSO_2.jpg}
\end{frame}

\begin{frame}{GA Results 1}
GA has much better performance on the 4 point designs.
\includegraphics[scale=0.28]{quadplots/GA_1.jpg}
\includegraphics[scale=0.28]{quadplots/GA_2.jpg}\\
In the case when $\beta = (2,0,-0.1)^T$, the design produced by GA is not still not quite optimal. Parameter tuning?
\end{frame}

\begin{frame}{GA Results 2}
GA can also find the optimal 3 point designs.
\includegraphics[scale=0.28]{quadplots/GA_3.jpg}
\includegraphics[scale=0.28]{quadplots/GA_4.jpg}
\end{frame}

\begin{frame}{Other Designs}
\includegraphics[scale=0.28]{quadplots/example_1.jpg}
\includegraphics[scale=0.28]{quadplots/stackex.jpg}
\end{frame}

\begin{frame}{Optimal designs for quadratic model on any interval.}
The previous results were all for the case where $x \in (-\infty, \infty)$. What if the design interval was one of the following?
\begin{itemize}
\item $[0, \infty)$
\item $[-1, 1]$
\item $[a,b]$ where $a,b$ are larger than the optimal design points on $(-\infty, \infty)$.
\item $[0,1]$
\end{itemize}
I will use GA to find these optimal designs as GA seems to have better performance compared to PSO.
\end{frame}

\begin{frame}{$[0, \infty)$}
Strictly positive designs have 3 support points.\\
\includegraphics[scale=0.2]{quadplots/positive_1.jpg}
\includegraphics[scale=0.2]{quadplots/positive_2.jpg}\\
\includegraphics[scale=0.2]{quadplots/positive_3.jpg}
\includegraphics[scale=0.2]{quadplots/positive_4.jpg}
\end{frame}

\begin{frame}{$[-1, 1]$}
When beta = (2, 0,-4), the global optimal design points are already within the interval, so we get the 4 point design.\\
\includegraphics[scale=0.2]{quadplots/11_1.jpg}
\includegraphics[scale=0.2]{quadplots/11_2.jpg}\\
\includegraphics[scale=0.2]{quadplots/11_3.jpg}
\includegraphics[scale=0.2]{quadplots/11_4.jpg}
\end{frame}

\begin{frame}{$[6,16]$}
beta=(2,0,-4) has issues. 1 point at 6 is close. Same with beta=(-2,0,-4). The two other nominal values are better.
\includegraphics[scale=0.28]{quadplots/616_1.jpg}
\includegraphics[scale=0.28]{quadplots/616_3.jpg}
\end{frame}

\begin{frame}{$[0,1]$}
Using the unit interval produces 3 point designs for out set of nominal values.\\
\includegraphics[scale=0.2]{quadplots/01_1.jpg}
\includegraphics[scale=0.2]{quadplots/01_2.jpg}\\
\includegraphics[scale=0.2]{quadplots/01_3.jpg}
\includegraphics[scale=0.2]{quadplots/01_4.jpg}
\end{frame}

\begin{frame}{beta = (2, 0, -0.1)}
\includegraphics[scale=0.18]{quadplots/GA_1.jpg}
\includegraphics[scale=0.18]{quadplots/positive_1.jpg}
\includegraphics[scale=0.18]{quadplots/11_1.jpg}
\includegraphics[scale=0.18]{quadplots/616_1.jpg}
\includegraphics[scale=0.18]{quadplots/01_1.jpg}

\end{frame}

\begin{frame}{beta= (2,0,-4)}
\includegraphics[scale=0.18]{quadplots/GA_2.jpg}
\includegraphics[scale=0.18]{quadplots/positive_2.jpg}
\includegraphics[scale=0.18]{quadplots/11_2.jpg}
%\includegraphics[scale=0.18]{quadplots/616_2.jpg}
\includegraphics[scale=0.18]{quadplots/01_2.jpg}
\end{frame}

\begin{frame}{beta=(-2,0,-0.1)}
\includegraphics[scale=0.18]{quadplots/GA_3.jpg}
\includegraphics[scale=0.18]{quadplots/positive_3.jpg}
\includegraphics[scale=0.18]{quadplots/11_3.jpg}
\includegraphics[scale=0.18]{quadplots/616_3.jpg}
\includegraphics[scale=0.18]{quadplots/01_3.jpg}
\end{frame}

\begin{frame}{beta=(-2,0,-4)}
\includegraphics[scale=0.18]{quadplots/GA_4.jpg}
\includegraphics[scale=0.18]{quadplots/positive_4.jpg}
\includegraphics[scale=0.18]{quadplots/11_4.jpg}
%\includegraphics[scale=0.18]{quadplots/616_4.jpg}
\includegraphics[scale=0.18]{quadplots/01_4.jpg}
\end{frame}

\begin{frame}{Conclusions}
\begin{itemize}
\item My code is able to replicate previous results for locally D-optimal designs.
\item Also is able to find designs on user-specified  design intervals.
\item Code can easily be extended to any degree polynomial.
\item Future work: fractional polynomials.
\item Demonstrate PlatEMO code.
\end{itemize}
\end{frame}

\begin{frame}{References}
\begin{itemize}
\item Paola Sebastiani, Raffaella Settimi (1997),
\textit{A note on D-optimal designs for a logistic regression model},
Journal of Statistical Planning and Inference
\item Shwetank Lall, Seema Jaggi, Eldho Varghese, Cini Varghese \& Arpan
Bhowmik (2018): \textit{An algorithmic approach to construct D-optimal saturated designs for logistic
model}, Journal of Statistical Computation and Simulation
\item Thomas Mathew, Bikas Kumar Sinha (2001),
\textit{Optimal designs for binary data under logistic regression},
Journal of Statistical Planning and Inference
\item Ellinor Fornius (2005), \textit{D-optimal Designs for Quadratic Logistic Regression Models}
\end{itemize}
\end{frame}
%\begin{frame}{Results from Mathew, Sinha}
%Mathew, Sinha (2001) derive several optimal designs for the 2 parameter logistic model.
%\begin{itemize}
%\item D-optimal: equal weights at $\frac{\pm 1.5434 - \beta_0}{\beta_1}$
%\item A-optimal: found best 2 point design.
%\end{itemize}
%Not addressed: A-optimal design within class of all designs and E-optimal designs. Also, no sensitivity plots to confirm optimality graphically.
%
%\end{frame}
%
%\begin{frame}{Results from Lall et al.}
%Lall et. al (2018) uses a modified Fedorov algorithm to construct saturated D-optimal designs for several 1 and 2 variable logistic models on the design interval $[-1,1]$. The following table shows their designs for the 1 variable model.
%\begin{center}
%\begin{tabular}{|c|c|c|c|}
%\hline 
% $\beta_0$ & $\beta_1$ & $x_1$ & $x_2$ \\ 
%\hline 
% 0.1 & 0.5 & -1 & 1 \\ 
%\hline 
% 1 & 1 & -1 & 1 \\ 
%\hline 
% 1 & 4 & -0.636 & 0.136 \\ 
%\hline 
%\end{tabular} 
%\end{center}
%Since these are saturated designs, both support points have equal weights.
%
%Their algorithm relies on a discretized design space, which makes it difficult to apply to global optimality.
%
%\end{frame}
%
%\begin{frame}{Results from Sebastiani}
%Sebastiani (1992) derives D-optimal designs for the 1 variable model for an unbounded, bounded design interval, and for a design interval bounded at a single end.
%\begin{itemize}
%\item Designs for bounded intervals are modified versions of the global optimum.
%\item If the ideal design point is outside of the bound, the bound will be one of the design points.
%\end{itemize}
%
%Results for $\beta_0 = 0$, $\beta_1 = 1$:
%\begin{center}
%\begin{tabular}{|c|c|c|c|c|c|}
%\hline 
%  & $[-\infty, \infty]$  & $[-1, \infty]$ & $[-\infty,1]$ & $[-1,1]$ & $[10, 20]$\\ 
%\hline 
%$x_1$ & -1.543 & -1 & -1.796 & -1  & 10.00\\ 
%\hline 
%$x_2$ & 1.543 & 1.796 & 1 & 1 & 12.00\\ 
%\hline 
%\end{tabular} 
%\end{center}
%All designs are equally weighted.
%
%\end{frame}
%
%\begin{frame}{Results from Fornius}
%Fornious (2005) derives D-optimal designs for the quadratic model with $\eta_i = \beta_0 + \beta_1 x_i + \beta_2 x_i^2$. Designs have either 3 or 4 support points depending on shape of response.
%
%\begin{tabular}{|c|c|}
%\hline 
% Nominal values & Design \\ 
%\hline 
% $2,0,-0.1$ & 
%$\begin{bmatrix}
%-5.7185 & -2.7017  & 2.7017 & 5.7185\\
%0.3138 & 0.1862 & 0.1862 & 0.3138
%\end{bmatrix}$  \\ 
%\hline 
% $2,0,-4$ & $\begin{bmatrix}
%-0.9042 & -0.4272  & 0.4272 & 0.9042\\
%0.3138 & 0.1862 & 0.1862 & 0.3138
%\end{bmatrix}$   \\ 
%\hline 
% $-2,0,-0.1$ & 
%$\begin{bmatrix}
%-3.9819 & 0  & 3.9819\\
%1/3 & 1/3 & 1/3
%\end{bmatrix}$   \\ 
%\hline 
% $-2,0,-4$ & 
%$\begin{bmatrix}
%-0.6296 & 0  & 0.6296\\
%1/3 & 1/3 & 1/3
%\end{bmatrix}$  \\ 
%\hline 
%\end{tabular} \\
%What happens when $\beta_1 \neq 0$? $\implies$ just shifts solution What about on different design intervals?
%\end{frame}

\end{document}

### 2021-06-11 project.m

@@ -1,0 +1,81 @@
+% find locally optimal designs for logistic model
+
+% design options
+beta = [2, 0, -4]; % nominal values, enter 3 values to add squared term
+lower = -Inf; % design interval
+upper = Inf;
+numpts = 4; % number of design ponts
+
+% algorithm options
+method = @GA; % algorithm to use
+iterations = 100; % number of iterations
+swarm = 1000; % size of the swarm
+proC = 1; % prob. of crossover
+disC = 20; % distr. index of cross-over: variance around parents
+proM = 1; % prob. of mutation
+disM = 20; % distr. index of mutation: variance of mutation
+
+% work in progress
+% matrix singularity gives issues with A and E opt.
+opt = 'D';
+
+% function evaluations
+maxFE = iterations * swarm; 
+
+% guess the interval the 2 design points are in
+% this doesn't matter so much since the repair function will move the
+% points automatically
+init_low = -5; 
+init_up = 5;
+init = @(n, d) [unifrnd(init_low, init_up, n/2, numpts),...
+    unifrnd(0, 1, n/2, numpts)]; % init swarm in interval
+
+% create upper and lower bounds for design points and weights
+% length of these vectors determine the number of design points
+% need an equal number of weights
+upperbounds = [upper*ones([1,numpts]), Inf*ones([1,numpts])];
+lowerbounds = [lower*ones([1,numpts]), zeros(1,numpts)];
+
+% find the optimal design
+[Dec,Obj,Con] = platemo('objFcn', @(x,d) logistic(x, beta, opt),...
+    'lower', lowerbounds, 'upper', upperbounds, ...
+    'algorithm', {method, proC, disC, proM, disM}, ...
+    'decFcn', @(x, d) sum_constraints(x, upper, lower),...
+    'maxFE', maxFE, ...
+    'N', swarm, ...
+    'initFcn', init);
+
+% find the optimal design
+out = sortrows([Dec, Obj], numpts*2 + 1); % sort by objective function
+xi = out(1, :); % first row has the point with the lowest objective value
+disp("Design found:")
+disp(xi(1:numpts))
+disp(xi(numpts+1:end-1))
+fprintf("Objective value: %f\n", xi(end));
+
+
+% check if design points are optimal
+fprintf("Sensitivity function values:\n");
+for i = 1:numpts
[…]

### 2021-06-11 src/ch_logistic.m

@@ -1,0 +1,83 @@
+% sensitivity function for two parameter logistic model
+function y = ch_logistic(z, vars, beta, opt)
+    
+    numpts = length(vars)/2;
+    x = vars(1:numpts);
+    w = vars(numpts+1:end);
+    
+%     sigma = exp(beta(1) + beta(2)*xvars)./(1 + exp(beta(1) + beta(2)*xvars)).^2;
+%     g = exp(beta(1) + beta(2)*x)/(1+exp(beta(1) + beta(2)*x))^2;
+
+     % compute eta
+    lbeta = length(beta);
+    if lbeta == 2
+       eta = beta(1) + beta(2) * x;
+       etaz = beta(1) + beta(2) * z;
+    elseif lbeta == 3
+        eta = beta(1) + beta(2) * x + beta(3) * x.^2;
+        etaz = beta(1) + beta(2) * z + beta(3) * z^2;
+    else
+        disp("Polynomial not supported")
+        return
+    end
+    
+    % weight functions
+    % need to update for higher degree poly
+    sigma = exp(eta)./(1 + exp(eta)).^2;
+    g = exp(etaz)/(1+exp(etaz))^2;
+    
+    % Design matrix
+    M = 0;
+    for i = 1:numpts
+        if lbeta == 3
+            M_i = w(i)*sigma(i) .* [1 x(i) x(i)^2; x(i) x(i)^2 x(i)^3;...
+                x(i)^2 x(i)^3 x(i)^4];
+        else 
+            M_i = [w(i)*sigma(i) w(i)*sigma(i)*x(i); ...
+            w(i)*sigma(i)*x(i) w(i)*sigma(i)*x(i)^2];
+        end
+        M = M + M_i;
+    end
+    
+    %detM = det(M);
+    
+    Minv = inv(M);
+    
+    % switch optimality
+    if (opt == 'D')
+        if lbeta == 3
+            y = g * [1 z z^2] * Minv * [1;z;z^2] - 3;
+        else
+            y = g * [1 z] * Minv * [1;z] - 2;
+        end
+        
+    elseif (opt == 'A')
+        if lbeta == 3
+            y = g * [1 z z^2] * inv(M + 1e-10 * eye(2))^2 * [1;z;z^2] - trace(Minv);
+        else
+           y = g * [1 z] * inv(M + 1e-10 * eye(2))^2 * [1;z] - trace(Minv);
+        end
+        
+    elseif (opt == 'E')
+        
+        [V, D] = eig(M); 
[…]

### 2021-06-11 src/logistic.m

@@ -1,0 +1,55 @@
+% more general logistic objective
+function obj = logistic(vars, beta, opt)
+
+    % Figure out how many design points
+    numpts = length(vars)/2;
+    
+    % separate out design points and weights
+    x = vars(1:numpts);
+    w = vars(numpts+1:end);
+    
+    % compute eta
+    lbeta = length(beta);
+    if lbeta == 2
+       eta = beta(1) + beta(2) * x;
+    elseif lbeta == 3
+        eta = beta(1) + beta(2) * x + beta(3) * x.^2;
+    else
+        disp("Polnomial not supported")
+        return
+    end
+    
+    % weight function
+    % need to update for higher degree poly
+    sigma = exp(eta)./(1 + exp(eta)).^2;
+    
+    % Design matrix
+    M = 0;
+    for i = 1:numpts
+        if lbeta == 3
+            M_i = w(i)*sigma(i) * [1 x(i) x(i)^2; x(i) x(i)^2 x(i)^3;...
+                x(i)^2 x(i)^3 x(i)^4];
+        else 
+            M_i = [w(i)*sigma(i) w(i)*sigma(i)*x(i); ...
+            w(i)*sigma(i)*x(i) w(i)*sigma(i)*x(i)^2];
+        end
+        
+        M = M + M_i;
+    end
+    
+    
+    % switch between optimality objectives
+    if (opt == 'D')
+        obj = -log(det(M));
+        
+    elseif (opt == 'A')
+        
+        obj = trace(inv(M + eye(2)));
+        
+    elseif (opt == 'E') % experiment with E-opt
+        obj = -min(eig(M));
+    else % default at D optimality
+        obj = -log(det(M));
+    end
+
+end

### 2021-06-11 src/plot_logistic.m

@@ -1,0 +1,21 @@
+function plot_logistic(xi, beta, opt, low, up)
+    
+    x = linspace(low, up, 100);
+    y = arrayfun(@(x) ch_logistic(x, xi, beta, opt), x);
+    plot(x,y)
+    xlim([low, up])
+    yline(0, '--')
+    n = length(xi)/2;
+    for i = 1:n
+        xline(xi(i), "--r", sprintf('x=%0.3f\n w=%0.3f', xi(i), xi(n+i)),...
+            'LabelVerticalAlignment', 'bottom',...
+            'LabelHorizontalAlignment', 'center');
+    end
+    if length(beta) == 2
+        title(sprintf('%s = (%0.2f, %0.2f)', '\beta', beta(1), beta(2)))
+    elseif length(beta) == 3
+        title(sprintf('%s = (%0.2f, %0.2f, %0.2f)', '\beta', beta(1), beta(2), beta(3)))
+    end
+    
+    return
+end

### 2021-06-11 src/sum_constraints.m

@@ -1,0 +1,20 @@
+% repair function to make weights sum to 1
+% TODO: simplex design space
+function vars = sum_constraints(vars, upper, lower)
+    
+
+    numpts = length(vars)/2;
+    x = vars(1:numpts);
+    w = vars(numpts+1:end);
+    
+    % scale weights
+    S = sum(w);
+    vars(numpts+1:end) = abs(w./S);
+    
+    % keep the bounding for the x's
+    % this code is copied from UserProblem.m in PlatEMO
+    vars(1:numpts) = max(min(x', repmat(upper, numpts, 1)), ...
+        repmat(lower, numpts, 1))';
+    
+return
+end

### 2021-06-11 paper/paper.tex
finished crossoever section

@@
 
-The following paragraphs detail the implementation of the genetic algorithm in PlatEMO. Let $N$ be the population size and let $D$ be the number of variables in the optimization problem.
+The following paragraphs detail the implementation of the genetic algorithm in PlatEMO.
+Let $N$ be the population size and let $D$ be the number of variables in the optimization problem.
+For simplicity, the description will assume $D=1$, but the process is eaily generalized to an arbritrary number of variables.
 
@@
 \end{flushleft}
-Once the mating pool has been selected, the two parents enter the crossover stage where genetic information is exchanged. This is done using simulated binary crossover (Deb et. al. 2007).
+Once the mating pool has been selected, the two parents enter the crossover stage where genetic information is exchanged.
+This is done using the simulated binary crossover procedure (Deb et.
+al.
+2007).
+This procedure will generate a probability distribution from which the child characteristics are drawn from.
+
+First, $\mu$ is drawn from the uniform distribution on $(0,1)$. If $u \leq 0.5$, then set $\beta = 2\mu^{1/(d_c+1)}$. If $\mu > 0.5$, then set $\beta = (2-2\mu)^{-1/(d_c+1)}$.
+
+The $d_c$ parameter is the distribution index of the simulated binary crossover and is related to the variance of the child solutions from the parents.
+If $d_c$ is large, then the children will likely be close to the parents.
+If $d_c$ is small, then the child solutions are more likely to be farther from the parents.
+
+The $\beta$ calculated in this way is a spread factor defined as 
+$$
+\beta = \left| \frac{C_2 - C_1}{P_2 - P_1} \right|
+$$
+where $C_1, C_2$ are child values and $P_1, P_2$ are the parent values. Therefore $\beta$ measures the distance between children relative to the distance between parents.
+
+Once $\beta$ has been found, the child solutions may be computed as
+$$
+C_1 = 0.5\left[(1+\beta)P_1 + (1-\beta)P_2 \right]
+$$
+and 
+$$
+C_2 = 0.5\left[(1-\beta)P_1 + (1+\beta)P_2 \right]
+$$
+
+Finally, this update is applied from probability $p_c$.
+In the implementation used by PlatEMO, additional steps are to set an expected number of crossovers for each pair of parents.
+On average, we should see $0.25 p_cD$ crossovers.
 
@@
 \end{flushleft}
+After crossover, each child solution is randomly perturbed.
+This helps the algorithm to better explore the objective space and hopefully escape local optima.
+This is acheived using a polnomial probabily distribution with mean at the current value and with variance equal to a function of $d_M$, the distribution index of mutation.
 

### 2021-06-16 paper/paper.tex
wrote conclusion

@@
 
-\subsection{Particle Swarm}
 \subsection{Genetic Algorithm}
@@
 \end{flushleft}
+Once the crossover and mutation operations take place, there will be a population of size $2N$.
+The objective function is evaluated for each of these solutions and the best $N$ are chosen to move on to the next generation.
 
@@
 We used the built-in algorithm ``GA'' with default parameters ($p_C = 1, d_C = 20, p_M = 1, d_M = 20$).
-We found that running the algorithm for 100000 function evaluations with a swarm size of 1000 was able produce close to optimal designs for every design we tried.
+We found that running the algorithm for 100000 function evaluations with a swarm size of 1000 was able produce near optimal designs for every design we tried.
 
 When the optimal number of design points was unknown, we ran the algorithm once with 10 design points and then made a guess at the correct number of design points.
-The optimal number of design points was easy to find after seeing the sensitivity plot.
+The optimal number of design points was easy to guess after seeing the sensitivity plot.
 
@@
 \section{Conclusions}
+In this paper, we find locally D-optimal designs for linear and quadratic logistic regression using a genetic algorithm.
+We confirm existing analytical results for the linear model on any design interval.
+Additionally, we find a new design in the case when the design interval does not include the original optimal design points on the unbounded interval.
+Finally, we confirm previous results for a quadratic model and find new designs for a bounded design interval.
+
+ Previous results depend on a variety of analytic and numeric methods to find the same designs.
+The advantage of our approach is that the genetic algorithm is a single method for finding optimal designs.
+This makes it easy for an investigator to change design parameters and find optimal designs for multiple scenarios.
+
+Our approach can be extended to other design problems for the logistic regression model.
+This paper considered locally D-optimal designs, but future work may consider A, E, and G optimal designs.
+The logit link function may also be exchanged for other link functions.
+An extension to fractional polynomials (Royston \& Altman, 1994) may also be possible.
+
 \section{References}

### 2021-06-18 paper/paper.tex
started work on intro

@@
 \section{Introduction}
+Logistic regression is a popular model for linking a binary outcome to one or more covariates.
+For the case of a single covariate, the model links a linear function of the covariate to the mean of the Bernouolli distribution.
+The resulting model can then be used to model the probability of an event occuring given a certain value of the covariate.
+This model has many extensions, including the addition of a squared term to the linear predictor.
+
+Before using logistic regression, it is important to design the experiment with the model in mind.
+For a fixed sample size $N$, we may wish to measure the covariate and outcome at certain points such that optimal model performance is obtained.
+Additionally, we should also consider how many of the covariate measurements to take at each point in the design space.
+Therefore, we want to find an optimal design $\xi$ such that the model achieves optimal performance according to some optimality criterion.
+A D-optimal design minimizes the volume of the confidence ellipsoid and is a common approach to this problem.
+For non-linear models, such as logistic regression, the D-optimal design depends on the parameters, so we looks for a locally optimal design given the paramters.
+These nomial values may be obtained from an intial analysis or from previous results.
+
+There are many papers on optimal designs for logistic regression.
+The locally D-optiaml design for the lienar predictor $\eta = \beta_0 + \beta_1 x$ can be found analytically.
+Ford (1972) proposed a D-optimal design on an unbounded interval \textbf{CHECH}.
+Sebastinani and Settimi (1992) derive D-optimal designs on several bounded design intervals.
+Numerical appproaches have also been treid.
+For example, Lall et.
+al (2018) find D-optimal designs on using a modified Fedorov algorithm.
+
+For the quadratic model with linear predictor $\eta = \beta_0 + \beta_1 + \beta_2x^2$, there have been no analytic results.
+Fornious (2005) found a D-optimal design for the quadratic model numerically.
+
+Although some optimal designs for simple models have been obtained analytically, the optimal designs for more complicated models may be difficult or impossible to obtain in this way.
+Traditional numerical optimization algorithms may help, they may also be complicated to implement.
+In both apporaches, minor changes to design may have major consequences in the ability to obtain the opitmal design.
+
 \section{Preliminaries}

### 2021-06-18 paper/paper.tex
finished introduction

@@
 Sebastinani and Settimi (1992) derive D-optimal designs on several bounded design intervals.
-Numerical appproaches have also been treid.
+Numerical appproaches have also been tried.
 For example, Lall et.
@@
 For the quadratic model with linear predictor $\eta = \beta_0 + \beta_1 + \beta_2x^2$, there have been no analytic results.
-Fornious (2005) found a D-optimal design for the quadratic model numerically.
+Fornious (2005) found locally D-optimal designs numerically for the quadratic model on an unbounded interval.
+Results for a bounded interval are unknown.
 
-Although some optimal designs for simple models have been obtained analytically, the optimal designs for more complicated models may be difficult or impossible to obtain in this way.
-Traditional numerical optimization algorithms may help, they may also be complicated to implement.
-In both apporaches, minor changes to design may have major consequences in the ability to obtain the opitmal design.
+For more complicated models, such as the quadratic logistic model, it can ber difficult an optimal design analytically.
+Numerical approaches such as by Fornious also rely on significant analytical work to find the optimal value of the objective function.
+Small changes to the model or to the design interval might result in a completely new problem with new analytical challenges.
+
+Another approach is to use a general purpose optimization algorithm to find the optimal designs.
+However, some algorithms are better suited for the task.
+For example, it can difficult to take the gradient of the objective function for many optimal design problems.
+This is more difficult by the fact that changing the model will result in a new gradient that has to be derived.
+Therefore, it is better to use a gradient-free algorithm for these problems.
+Another problem that arises is local optima.
+Thus, we will want to use an optimization algorithm that can ``explore" the optimization space without getting stuck in local optima.
+
+Nature-inspired metaheuristics are a family of algorithms that are gradient free and have mechanisms for escaping local optima.
+Examples include particle swarm optimization (Kennedy, Eberhart,  1995), the artificial bee colony algorithm (Karaboğa, 2005), and differential evolution (Storn, Price, 1997).
+These algorithms mimic the behavior of animals or nature processes to find a global optimum.
+
+In this paper, we use the genetic algorithm (Holland, 1975) which is inspired by natural selection.
+The algorithm starts with an initial population of solution vectors and then combines and mutates the solutions to hopefully produce more solutions that have a better objective function value.
+The genetic algorithm has many different implementations and is widely applied to solve optimization problems.
+In this paper, we use the genetic algorithm to find locally D-optimal designs for several logistic regression models.
 

### 2021-06-18 paper/paper.tex
finished references

@@
 Logistic regression is a popular model for linking a binary outcome to one or more covariates.
-For the case of a single covariate, the model links a linear function of the covariate to the mean of the Bernouolli distribution.
-The resulting model can then be used to model the probability of an event occuring given a certain value of the covariate.
+For the case of a single covariate, the model links a linear function of the covariate to the mean of the Bernoulli distribution.
+The resulting model can then be used to model the probability of an event occurring given a certain value of the covariate.
 This model has many extensions, including the addition of a squared term to the linear predictor.
@@
 A D-optimal design minimizes the volume of the confidence ellipsoid and is a common approach to this problem.
-For non-linear models, such as logistic regression, the D-optimal design depends on the parameters, so we looks for a locally optimal design given the paramters.
-These nomial values may be obtained from an intial analysis or from previous results.
+For non-linear models, such as logistic regression, the D-optimal design depends on the parameters, so we looks for a locally optimal design given the parameters.
+These nominal values may be obtained from an initial analysis or from previous results.
 
 There are many papers on optimal designs for logistic regression.
-The locally D-optiaml design for the lienar predictor $\eta = \beta_0 + \beta_1 x$ can be found analytically.
-Ford (1972) proposed a D-optimal design on an unbounded interval \textbf{CHECH}.
-Sebastinani and Settimi (1992) derive D-optimal designs on several bounded design intervals.
+The locally D-optimal design for the linear predictor $\eta = \beta_0 + \beta_1 x$ can be found analytically.
+Ford et.
+al.
+[2] proposed a D-optimal design on an unbounded interval.
+Sebastinani and Settimi[11] derive D-optimal designs on several bounded design intervals.
 Numerical appproaches have also been tried.
 For example, Lall et.
-al (2018) find D-optimal designs on using a modified Fedorov algorithm.
+al [7] find D-optimal designs on using a modified Fedorov algorithm.
 
 For the quadratic model with linear predictor $\eta = \beta_0 + \beta_1 + \beta_2x^2$, there have been no analytic results.
-Fornious (2005) found locally D-optimal designs numerically for the quadratic model on an unbounded interval.
+Fornius [3] found locally D-optimal designs numerically for the quadratic model on an unbounded interval.
 Results for a bounded interval are unknown.
 
-For more complicated models, such as the quadratic logistic model, it can ber difficult an optimal design analytically.
-Numerical approaches such as by Fornious also rely on significant analytical work to find the optimal value of the objective function.
+For more complicated models, such as the quadratic logistic model, it can be difficult an optimal design analytically.
+Numerical approaches such as by Fornius also rely on significant analytical work to find the optimal value of the objective function.
 Small changes to the model or to the design interval might result in a completely new problem with new analytical challenges.
@@
 Nature-inspired metaheuristics are a family of algorithms that are gradient free and have mechanisms for escaping local optima.
-Examples include particle swarm optimization (Kennedy, Eberhart,  1995), the artificial bee colony algorithm (Karaboğa, 2005), and differential evolution (Storn, Price, 1997).
+Examples include particle swarm optimization [6], the artificial bee colony algorithm [5], and differential evolution [9].
 These algorithms mimic the behavior of animals or nature processes to find a global optimum.
 
-In this paper, we use the genetic algorithm (Holland, 1975) which is inspired by natural selection.
+In this paper, we use the genetic algorithm [4] which is inspired by natural selection.
 The algorithm starts with an initial population of solution vectors and then combines and mutates the solutions to hopefully produce more solutions that have a better objective function value.
@@
 
-The genetic algorithm (Holland, 1975) is a nature-inspired metaheuristic optimization algorithm.
+The genetic algorithm 43] is a nature-inspired metaheuristic optimization algorithm.
 The algorithm mimics natural selection with the intended goal of ``evolving" a population of solution vectors towards an optimal value.
@@
 Let $N$ be the population size and let $D$ be the number of variables in the optimization problem.
-For simplicity, the description will assume $D=1$, but the process is eaily generalized to an arbritrary number of variables.
+For simplicity, the description will assume $D=1$, but the process can be generalized to an arbitrary number of variables.
 
@@
 \end{flushleft}
-The first step in the genetiec algorithm is to select the mating pool via tournament selection.
+The first step in the genetic algorithm is to select the mating pool via tournament selection.
 $N$ paris are sampled with replacement from the population.
-The objective function values for eahc of these pairs is combared and the most opitmal of the pair is selected.
+The objective function values for each of these pairs is compared and the most optimal of the pair is selected.
 
@@
 Once the mating pool has been selected, the two parents enter the crossover stage where genetic information is exchanged.
-This is done using the simulated binary crossover procedure (Deb et.
-al.
-2007).
+This is done using the simulated binary crossover procedure [1].
 This procedure will generate a probability distribution from which the child characteristics are drawn from.
@@
 This helps the algorithm to better explore the objective space and hopefully escape local optima.
-This is acheived using a polnomial probabily distribution with mean at the current value and with variance equal to a function of $d_M$, the distribution index of mutation.
+This is achieved using a polynomial probability distribution with mean at the current value and with variance equal to a function of $d_M$, the distribution index of mutation.
 
@@
 Therefore, at least one of the two design points will likely be one of the bounds of the interval.
-Theoretical justifications for this phenomenon can be found in Sebastiani.
+Theoretical justifications for this phenomenon can be found in Sebastiani [11].
 
@@
 There are no analytic results for locally D-optimal designs when $\eta_i = \beta_0 + \beta_1 x_i + \beta_2 x_i^2$.
-However, Fornious was able to find locally D-optimal designs numerically for several nominal $\beta$.
+However, Fornius [3] was able to find locally D-optimal designs numerically for several nominal $\beta$.
 If $\beta_1 \neq 0$, then the design shifts along the design space.
@@
 The logit link function may also be exchanged for other link functions.
-An extension to fractional polynomials (Royston \& Altman, 1994) may also be possible.
+An extension to fractional polynomials [10] is also possible.
+
+\newpage
 
 \section{References}
+\begin{enumerate}
+\item K. Deb, K. Sindhya, and T. Okabe, Self-adaptive simulated binary
+crossover for real-parameter optimization, Proceedings of the Annual
+Conference on Genetic and Evolutionary Computation, 2007, 1187-1194.
+
+\item Ford, I.
+, Torsney, B.
+and Wu, C.F.J.
+(1992) The use of a canonical form in the construction of locally optimal designs for non-linear problems.
+Journal of the Royal Statistical Society: Series B (Statistical Methodology), 54(2), pp.
+569-583.
+\item Fornius, E. (2007) D-optimal Designs for Quadratic Logistics Regression Models. International Journal of Statistical Sciences, Vol. 6, 2007, pp 269-303.
+
+\item Holland, John (1992). Adaptation in Natural and Artificial Systems. Cambridge, MA: MIT Press. ISBN 978-0262581110
+\item Karaboga, D. (2005). An Idea Based on Honey Bee Swarm for Numerical Optimization.
+\item J. Kennedy and R. Eberhart, "Particle swarm optimization," Proceedings of ICNN'95 - International Conference on Neural Networks, 1995, pp. 1942-1948 vol.4, doi: 10.1109/ICNN.1995.488968.
+\item  Shwetank Lall, Seema Jaggi, Eldho Varghese, Cini Varghese \& Arpan Bhowmik (2018) An algorithmic approach to construct D-optimal saturated designs for logistic model, Journal of Statistical Computation and Simulation, 88:6, 1191-1199, DOI: 10.1080/00949655.2018.1425689
+
+\item Thomas Mathew, Bikas Kumar Sinha,
+Optimal designs for binary data under logistic regression,
+Journal of Statistical Planning and Inference,
+Volume 93, Issues 1–2,
+2001,
+Pages 295-307,
+ISSN 0378-3758,
+https://doi.org/10.1016/S0378-3758(00)00173-7.
+
+\item Storn, R., Price, K.
+Differential Evolution – A Simple and Efficient Heuristic for global Optimization over Continuous Spaces.
+Journal of Global Optimization 11, 341–359 (1997).
+https://doi.org/10.1023/A:1008202821328
+
+\item Royston P, Altman DG (1994) Regression using fractional polynomials of continuous covariates: parsimonious parametric modelling (with discussion). Appl Stat 43(3):429–467
+
+\item Paola Sebastiani, Raffaella Settimi,
+A note on D-optimal designs for a logistic regression model,
+Journal of Statistical Planning and Inference,
+Volume 59, Issue 2, 
+1997, 
+https://doi.org/10.1016/S0378-3758(96)00111-5.
+
+\end{enumerate}
+
+\newpage
+
 \section{Appendix}

### 2021-06-18 paper/paper.tex
editing pass

@@
 \usepackage{graphicx}
+\usepackage{hyperref}
 
@@
 \section{Introduction}
-Logistic regression is a popular model for linking a binary outcome to one or more covariates.
-For the case of a single covariate, the model links a linear function of the covariate to the mean of the Bernoulli distribution.
-The resulting model can then be used to model the probability of an event occurring given a certain value of the covariate.
+Logistic regression is a popular approach for modeling the relationship between a binary outcome and one or more covariates.
+For the case of a single covariate, the model links a linear function of the covariate to the mean of a Bernoulli distribution.
+The resulting model can then be used to model the probability of an event occurring, given a certain value of the covariate.
 This model has many extensions, including the addition of a squared term to the linear predictor.
@@
 Before using logistic regression, it is important to design the experiment with the model in mind.
-For a fixed sample size $N$, we may wish to measure the covariate and outcome at certain points such that optimal model performance is obtained.
-Additionally, we should also consider how many of the covariate measurements to take at each point in the design space.
+For a fixed sample size $N$, we may wish to measure the covariate and outcome at certain points such that the model performs optimally.
 Therefore, we want to find an optimal design $\xi$ such that the model achieves optimal performance according to some optimality criterion.
-A D-optimal design minimizes the volume of the confidence ellipsoid and is a common approach to this problem.
-For non-linear models, such as logistic regression, the D-optimal design depends on the parameters, so we looks for a locally optimal design given the parameters.
-These nominal values may be obtained from an initial analysis or from previous results.
+The choice of opitimality criterion will depend on the goal of the analysis.
+A D-optimal design minimizes the volume of the confidence ellipsoid and thus is a popular choice of criterion.
 
 There are many papers on optimal designs for logistic regression.
-The locally D-optimal design for the linear predictor $\eta = \beta_0 + \beta_1 x$ can be found analytically.
+A few relevant papers relating to finding D-optimal designs for linear and quadratic predictors are reviewed here.
 Ford et.
@@
 
-For the quadratic model with linear predictor $\eta = \beta_0 + \beta_1 + \beta_2x^2$, there have been no analytic results.
+For the quadratic model with linear predictor $\eta = \beta_0 + \beta_1 + \beta_2x^2$, there are fewer results.
 Fornius [3] found locally D-optimal designs numerically for the quadratic model on an unbounded interval.
-Results for a bounded interval are unknown.
+The D-optimal designs for a bounded interval are unknown.
 
-For more complicated models, such as the quadratic logistic model, it can be difficult an optimal design analytically.
-Numerical approaches such as by Fornius also rely on significant analytical work to find the optimal value of the objective function.
+For more complicated models, such as the quadratic logistic model, it can be difficult to find an optimal design analytically.
+Numerical approaches such as by Fornius also rely on significant analytical work to derive a suitable objective function.
 Small changes to the model or to the design interval might result in a completely new problem with new analytical challenges.
@@
 Another approach is to use a general purpose optimization algorithm to find the optimal designs.
-However, some algorithms are better suited for the task.
+However, some algorithms are better suited for the task than others.
 For example, it can difficult to take the gradient of the objective function for many optimal design problems.
-This is more difficult by the fact that changing the model will result in a new gradient that has to be derived.
+This is made more difficult by the fact that changing the model often results in a new gradient.
 Therefore, it is better to use a gradient-free algorithm for these problems.
-Another problem that arises is local optima.
-Thus, we will want to use an optimization algorithm that can ``explore" the optimization space without getting stuck in local optima.
+Another common problem that arises in optimal design is local optima.
+We will want to use an optimization algorithm that can ``explore" the optimization space without getting stuck in the local optima.
 
@@
 The algorithm starts with an initial population of solution vectors and then combines and mutates the solutions to hopefully produce more solutions that have a better objective function value.
-The genetic algorithm has many different implementations and is widely applied to solve optimization problems.
-In this paper, we use the genetic algorithm to find locally D-optimal designs for several logistic regression models.
+The genetic algorithm has many different implementations and has been widely applied to solve optimization problems in science and engineering.
+In this paper, we use a genetic algorithm to find D-optimal designs for  logistic regression models on any design interval.
 
@@
 In this case, $\sum_{i=1}^k w_i = 1$ where $w_i = n_i/N$.
-
-\par A design using weights be written as
+To convert back to the original count format, we may have to round to the near integer.
+A design using weights be written as
 $$
@@
 
-The design $x_i$ implies a corresponding information matrix for the logistic regression model. Let
+The design $\xi$ implies a corresponding information matrix for the logistic regression model. Let
 $$
@@
 However, $\Psi(M)$ also depends on the parameters $\beta_0$, $\beta_1$, and $\beta_2$.
-This means that nominal values will need to be supplied for these parameters.
+The resulting design will depend on nominal values for these parameters, which may be obtained from a preliminary analysis or from previous results.
 Therefore, we may obtain a locally D-optimal design by first providing values for the parameters and then minimizing $\Psi(M)$.
 
-To check if a design is locally D-optimal, we calculate the sensitivity function
+To check if a design is locally D-optimal, we can calculate the sensitivity function
 $$
@@
 $$
-The equivalence theorem says a design is locally D-optimal if and only if $ch(x) \leq 0 $ for all $x$ in the design interval with equality at the design points.
+A design is locally D-optimal if and only if $ch(x) \leq 0 $ for all $x$ in the design interval with equality at the design points.
 
 \subsection{Genetic Algorithm}
-Finding a locally D-optimal design for the logistic model requires we optimize a function of several variables.
+Finding a locally D-optimal design for the logistic model involves a function of design points and their weights.
 This can be accomplished analytically for a linear $\eta$, but the complexity dramatically increases for a quadratic $\eta$ and requires a numerical solution.
@@
 
-The genetic algorithm 43] is a nature-inspired metaheuristic optimization algorithm.
+The genetic algorithm [4] is a nature-inspired metaheuristic optimization algorithm.
 The algorithm mimics natural selection with the intended goal of ``evolving" a population of solution vectors towards an optimal value.
-Like other nature-inspired optimization algorithms, the genetic algorithm has the advantages of being gradient free, able to accommodate constraints, and able to escape local optimums.
+Like other nature-inspired optimization algorithms, the genetic algorithm has the advantages of being gradient free, able to accommodate constraints, and able to escape local optima.
 
-There are many variations on the genetic algorithm, but these algorithms usually contain the following three steps that are run every iteration of the algorithm.
+There are many variations of the genetic algorithm, but these algorithms usually contain the following three steps that are run every iteration or ``generation" of the algorithm.
 First, the \textit{crossover} operation combines two parent solutions to make two child solutions.
 Second, the \textit{mutation} operation randomly perturbs the child solutions.
-Finally, \textit{fitness} is assessed and the solution vectors that have the best objective function values move on to the next generation.
+Finally, \textit{fitness} is assessed and the solution vectors that have the best objective function values survive to the next generation.
 
-The following paragraphs detail the implementation of the genetic algorithm in PlatEMO.
+The following paragraphs give an overview of  the implementation of the genetic algorithm used in this paper.
+We use the version from PlatEMO, a MATLAB platform for evolutionary computing.[12] Our code is available at \href{https://github.com/willgertsch/ODlogitpoly}{github.com/willgertsch/ODlogitpoly}.
+
 Let $N$ be the population size and let $D$ be the number of variables in the optimization problem.
@@
 \end{flushleft}
-The first step in the genetic algorithm is to select the mating pool via tournament selection.
-$N$ paris are sampled with replacement from the population.
-The objective function values for each of these pairs is compared and the most optimal of the pair is selected.
+The first step in the genetic algorithm is to select the mating pool via a tournament selection algorithm.
+First, $N$ pairs are sampled with replacement from the population, then the objective function values for each of these pairs are compared.
+Finally, the best solutions of the pairs are selected for crossover.
 
@@
 \end{flushleft}
-Once the mating pool has been selected, the two parents enter the crossover stage where genetic information is exchanged.
+Once the mating pool has been selected, the two parents enter the crossover stage where ``genetic" information is exchanged.
 This is done using the simulated binary crossover procedure [1].
@@
 
-First, $\mu$ is drawn from the uniform distribution on $(0,1)$. If $u \leq 0.5$, then set $\beta = 2\mu^{1/(d_c+1)}$. If $\mu > 0.5$, then set $\beta = (2-2\mu)^{-1/(d_c+1)}$.
-
+First, $\mu$ is drawn from the uniform distribution on $(0,1)$.
+If $u \leq 0.5$, then set $\beta = 2\mu^{1/(d_c+1)}$.
+If $\mu > 0.5$, then set $\beta = (2-2\mu)^{-1/(d_c+1)}$.
 The $d_c$ parameter is the distribution index of the simulated binary crossover and is related to the variance of the child solutions from the parents.
@@
 We found that running the algorithm for 100000 function evaluations with a swarm size of 1000 was able produce near optimal designs for every design we tried.
+We were able to confirm optimality by plotting the sensitivity function on the design interval.
 
@@
 The genetic algorithm can find designs on an arbitrary design interval $[a,b]$.
-In general, the optimal design will change from the design on $(-\infty, \infty)$ if any of the designs points is no longer included in the interval.
+In general, the optimal design will change from the optimal design on $(-\infty, \infty)$ if any of the designs points are no longer included in the interval.
 Therefore, at least one of the two design points will likely be one of the bounds of the interval.
@@
 
-Table 2 shows how the design interval affects the optimal design points.
+Table 2 shows the effect of the bounded interval on the optimal design points.
 For example, if the interval excludes the global optimal point -1.543, then the optimal design will have one point at the bound of the design interval closest to that point.
@@
 \subsection{Quadratic $\eta$ on $(-\infty, \infty)$}
-There are no analytic results for locally D-optimal designs when $\eta_i = \beta_0 + \beta_1 x_i + \beta_2 x_i^2$.
-However, Fornius [3] was able to find locally D-optimal designs numerically for several nominal $\beta$.
+Analytic results for locally D-optimal designs when $\eta_i = \beta_0 + \beta_1 x_i + \beta_2 x_i^2$ are much more challenging to find compared to the linear model.
+Fornius [3] was able to find locally D-optimal designs numerically for several nominal $\beta$.
 If $\beta_1 \neq 0$, then the design shifts along the design space.
-Therefore, the example nominal values have $\beta_1$ set at 0.
+Therefore, the example nominal values used by Fornius have $\beta_1$ set at 0 since the value of $\beta_1$ has less influence on the design compared to the other parameters.
 
@@
 
-Unfortunately, not all designs have a sensitivity plot that can easily plotted.
+Unfortunately, not all designs have a sensitivity plot that can easily be interpreted.
 The genetic algorithm still finds the design but it is difficult to judge from the plot whether the design is truly optimal.
@@
 
-\section{Applications}
 \section{Conclusions}
@@
 
- Previous results depend on a variety of analytic and numeric methods to find the same designs.
-The advantage of our approach is that the genetic algorithm is a single method for finding optimal designs.
+ Previous results depend on a variety of analytic and numeric methods to find the optimal designs.
+The advantage of our approach is that the genetic algorithm is a single method for finding optimal designs across several different situations.
 This makes it easy for an investigator to change design parameters and find optimal designs for multiple scenarios.
@@
 
+\item Ye Tian, Ran Cheng, Xingyi Zhang, and Yaochu Jin, PlatEMO: A MATLAB Platform for Evolutionary Multi-Objective Optimization [Educational Forum], IEEE Computational Intelligence Magazine, 2017, 12(4): 73-87".
+
 \end{enumerate}
 
-\newpage
-
-\section{Appendix}
 \end{document}
